The role of advection versus diffusion in intraventricular transport

### 2025-02-03 paper/main.tex
Results 6

\subsection*{Modification of ventricles geometry impacts protein distribution}
We simulate flow and transport in four alternative brain ventricles meshes:
one where the connection between the anterior and middle ventricles is constricted;
one where the connection between the middle and posterior ventricles is constricted;
one where the middle ventricle is shrunk in the lateral direction;
the combination of the above three modifications~(\Cref{fig:figure6_sim_results_compare_geometries}a).
We simulate CSF flow in the modified geometries with the baseline flow model and use
the resulting velocity field to simulate transport of Dendra2.

Constricting the connection between the anterior and middle ventricle leads to
slower transport to region of interest (ROI) 5 located in the anterior ventricle,
and a slightly higher mean concentration in ROIs 2--4 and 6, compared to the original geometry.
On the other hand, constricting the connection between the middle and the posterior ventricle
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5).
Faster transport to ROI 5 and slower transport to ROI 6 is also
the case for the geometry with a shrunk middle ventricle.
Additionally, for this configuration the transport towards the anterior region of
the middle ventricle (ROI 3) is slowed down.
For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricles (ROIs 2--4).
Both the anterior and the posterior ventricles (ROIs 5 and 6) have a lower mean concentration,
compared to the results for the original geometry.

@@

\section*{Discussion}
We have presented a computational model of cerebrospinal fluid (CSF) flow in brain ventricles of 2 days post-fertilization (2dpf) zebrafish embryo. By imposing tangential stresses and normal traction on the CSF flow, the simulated flow fields had similar features to the flow fields observed experimentally, consisting of vortex-structured flow and pulsatile flow. The CSF flow was used to model transport of a solute by an advection-diffusion equation. Qualitatively, simulations compared well to experiments of Dendra2 fluorescent protein photoconversion. Transport in cilia-mutant fish (\emph{schmalhans} mutants) and a CSF model without cilia, was compared to control experiments and the baseline CSF model, indicating a local effect of cilia motion on protein transport. Using the computational model of flow and transport, we further studied region-specific removal of cilia and ventricles geometry modification, and how these changes affected transport.

>>>>

\subsection*{Modification of ventricles geometry impacts protein distribution}
To assess the impact of ventricular morphologies on solute distribution,
we used four alternative brain ventricles meshes:
one where the connection between the anterior and middle ventricles is constricted;
one where the connection between the middle and posterior ventricles is constricted;
one where the middle ventricle is shrunk in the lateral direction;
and the combination of the above three modifications (\Cref{fig:figure6_sim_results_compare_geometries}a).
We simulate CSF flow in the modified geometries with the baseline flow model
and use the resulting velocity field to simulate transport of Dendra2 (\Cref{fig:figure6_sim_results_compare_geometries}b).

Constricting the connection between the anterior and middle ventricle leads to
slower transport to the anterior ventricle (ROI 5),
and a slightly higher mean concentration in ROIs 2--4 and 6, compared to the original geometry.
On the other hand, constricting the connection between the middle and the posterior ventricle
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5).
Faster transport to the anterior ventricle (ROI 5) and slower transport to the
posterior ventricle (ROI 6) is also the case for the geometry with a shrunk middle ventricle.
Additionally, for this configuration the transport upstream the flow fields towards the
ventral-posterior region of the middle ventricle (ROI 3) is slowed down.
For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have lower concentration.
Altogether our results identify that the geometry and width of the ventricular ducts matter
for solute transport and may explain why we observed larger differences
in the cilia-deficient model than in the experimental data in
\Cref{fig:figure4_compare_control_mutant}, which have different duct diameters.

@@

\section*{Discussion}
We have presented a computational model of cerebrospinal fluid (CSF) flow in brain ventricles of 2 days post-fertilization (2dpf) zebrafish embryo.
By imposing tangential traction and normal pressure on the CSF flow, the simulated flow fields had similar features to the flow fields observed experimentally, consisting of vortex-structured flow and pulsatile flow. The CSF flow was used to model transport of a solute by an advection-diffusion equation. Qualitatively, simulations compared well to experiments of Dendra2 fluorescent protein photoconversion. Transport in cilia-mutant fish (\emph{schmalhans} mutants) and a CSF model without cilia, was compared to control experiments and the baseline CSF model, indicating a local effect of cilia motion on protein transport. Using the computational model of flow and transport, we further studied region-specific removal of cilia and ventricles geometry modification, and how these changes affected transport.
